---
title: "301-all-the-data"
author: "Mac Campbell"
date: "October 28, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(RColorBrewer)
```

## PCA
Let's look at the test chrom PCA to see what we get.     
`bash $HOME/shernook/201-parallelize-analyses.sh $HOME/shernook/bamlists/bamlist156.bamlist $HOME/shernook/meta/test.chroms`     
(this is in outputs/300)   
`gunzip -c NC_037097.1-pca.beagle.gz | head -n 1 > test-chroms.beagle`      
`gunzip -c *.beagle.gz | grep -v "marker" >> test-chroms.beagle`     
`gzip test-chroms.beagle`
`python $HOME/pcangsd/pcangsd.py -beagle test-chroms.beagle.gz -o test-chroms -threads 10`
Parsing Beagle file      
Read 156 samples and 1897667 sites      

Estimating population allele frequencies      
EM (MAF) converged at iteration: 14     

Number of sites after MAF filtering (0.05): 1159472     

```{r}
#' @param samples character vector with the individuals IDs in the order in which
#' they were passed in the bamlist to angsd.
#' @param cov covariance matrix
covar2pcs <- function(samples, cov) {
  
  
  eig <- eigen(cov, symm = TRUE)
  PC <- as.data.frame(eig$vectors) %>%
    as_tibble() %>%
    setNames(sprintf("PC-%02d", 1:ncol(.)))
  
  samtib <- tibble(sample = samples)
  
  list(
    PCs = bind_cols(samtib, PC),
    eigevalues = eig$values
  )
}

```

```{r}
cov<-read_delim("outputs/300/test-chroms.pcacov", col_names=FALSE, delim=" ") %>% as.matrix()

load("meta/tot.rda")

pca <- covar2pcs(tot$`Sample ID`, cov)

pca_long <- pca$PCs %>%
  tidyr::gather(., key = "PC", "val", -sample)

# then expand a grid of the possible comparisons (ordered)
expg <- expand.grid(sample = pca$PCs$sample,
                    PCx = sprintf("PC-%02d", 1:6),
                    PCy = sprintf("PC-%02d", 1:6),
                    stringsAsFactors = FALSE) %>%
  tibble::as_tibble()

# then left join the pca results onto that
pca_pairs <- dplyr::left_join(expg, pca_long, by = c("sample", "PCx" = "PC")) %>%
  dplyr::rename(val_x = val) %>%
  dplyr::left_join(pca_long, by = c("sample", "PCy" = "PC")) %>%
  dplyr::rename(val_y = val)

pp_meta <- pca_pairs %>%   # just keep the first 6 PCs around
  left_join(., tot, by = c("sample" = "Sample ID")) 

npc <- 3
pp_meta2 <- pp_meta %>%
  filter( (PCx %in% sprintf("PC-%02d", 1:npc)) & 
            (PCy %in% sprintf("PC-%02d", 1:npc)) )

eig <- eigen(cov, symm = TRUE)
var<-eig$values/sum(eig$values)
cumvar<-cumsum(eig$values)/sum(eig$values)

```

```{r}
ggplot(pp_meta2) +
  geom_point(aes(x=val_x, y=val_y, shape=Run, color=Tributary)) +
  coord_fixed() +
  facet_grid (PCx ~ PCy) +
  theme_bw()
```

Kinda gross... Dropping Winter Run     

```{r}
ppsub <- pp_meta2 %>% filter(Run != "Winter")

ggplot(ppsub) +
  geom_point(aes(x=val_x, y=val_y, shape=Run, color=Tributary)) +
  coord_fixed() +
  facet_grid (PCx ~ PCy) +
  theme_bw()
```

Well! That's not very readable still.     

## Association
__1__ Early-Late    
`bash $HOME/shernook/202-do-asso.sh $HOME/shernook/bamlists/bamlist156.bamlist $HOME/shernook/meta/test.chroms $HOME/shernook/bamlists/bamlist156-early-late.pheno`    
(this is in outputs/300)     
```{r}
d1 <-  read_tsv("outputs/300/NC_037097.1-asso.lrt0.gz")
d16 <- read_tsv("outputs/300/NC_037112.1-asso.lrt0.gz")
d28 <- read_tsv("outputs/300/NC_037124.1-asso.lrt0.gz")
d34 <- read_tsv("outputs/300/NC_037130.1-asso.lrt0.gz")

data<-bind_rows(d1,d16,d28,d34) 

data <- data %>% mutate(log10p = -log10(dchisq(LRT, df = 1))) %>%
  filter(log10p > 1) %>% filter(log10p!="Inf")

data %>% arrange(-log10p) %>% head()
```

Manhattan Style!!!

```{r}
dd <- data %>% mutate(Index=1:n())

chroms<-dd %>% group_by(Chromosome) %>% mutate(Start=min(Index), Stop=max(Index)) %>% 
  select(Chromosome,Start,Stop) %>% unique() %>% 
  mutate(Position=round((Start+Stop)/2))

outliers <- dd %>% filter(log10p >=10)
#Make enough colors
nb.cols <- nrow(chroms)
mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(nb.cols)


ggplot(dd) +
  geom_point(data=dd, aes(x=Index, y=log10p, color=Chromosome), alpha=0.75, cex=0.5) +
  geom_point(data=outliers, aes(x=Index, y=log10p), color="black", cex=0.5, alpha=0.9) +
  geom_hline(yintercept = 6, lty=2, alpha=0.6) +
  geom_hline(yintercept = 10, lty=2, alpha=0.6) +
  theme_bw() +
  scale_x_continuous(breaks=chroms$Position, labels=chroms$Chromosome) +
  theme(axis.text.x = element_text(angle=45, hjust = 1, face="bold", size=6)) +
  theme(legend.title = element_blank()) +
  theme(panel.grid = element_blank()) +
  theme(legend.position = "none") +
  scale_color_manual(values=mycolors) +
  ylab("-log10(p)") +
  xlab("Chromosome") +
  ggtitle("Early vs Late Comparison - Combined Data") +
  theme(plot.title = element_text(hjust=0.5) )

ggsave("outputs/300/early-late-combined-data.jpeg")
```


__2__ Winter-Spring    
`bash $HOME/shernook/202-do-asso.sh $HOME/shernook/bamlists/bamlist156-winter-spring.bamlist $HOME/shernook/meta/test.chroms $HOME/shernook/bamlists/bamlist156-winter-spring.pheno`     
(this is in outputs/301)    
22 winter-run
61 spring-run

```{r}
d1 <-  read_tsv("outputs/301/NC_037097.1-asso.lrt0.gz")
d16 <- read_tsv("outputs/301/NC_037112.1-asso.lrt0.gz")
d28 <- read_tsv("outputs/301/NC_037124.1-asso.lrt0.gz")
d34 <- read_tsv("outputs/301/NC_037130.1-asso.lrt0.gz")

data<-bind_rows(d1,d16,d28,d34) 

data <- data %>% mutate(log10p = -log10(dchisq(LRT, df = 1))) %>%
  filter(log10p > 1) %>% filter(log10p!="Inf")

data %>% arrange(-log10p) %>% head()
```

```{r}
dd <- data %>% mutate(Index=1:n())

chroms<-dd %>% group_by(Chromosome) %>% mutate(Start=min(Index), Stop=max(Index)) %>% 
  select(Chromosome,Start,Stop) %>% unique() %>% 
  mutate(Position=round((Start+Stop)/2))

outliers <- dd %>% filter(log10p >=10)
#Make enough colors
nb.cols <- nrow(chroms)
mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(nb.cols)


ggplot(dd) +
  geom_point(data=dd, aes(x=Index, y=log10p, color=Chromosome), alpha=0.75, cex=0.5) +
  geom_point(data=outliers, aes(x=Index, y=log10p), color="black", cex=0.5, alpha=0.9) +
  geom_hline(yintercept = 6, lty=2, alpha=0.6) +
  geom_hline(yintercept = 10, lty=2, alpha=0.6) +
  theme_bw() +
  scale_x_continuous(breaks=chroms$Position, labels=chroms$Chromosome) +
  theme(axis.text.x = element_text(angle=45, hjust = 1, face="bold", size=6)) +
  theme(legend.title = element_blank()) +
  theme(panel.grid = element_blank()) +
  theme(legend.position = "none") +
  scale_color_manual(values=mycolors) +
  ylab("-log10(p)") +
  xlab("Chromosome") +
  ggtitle("Spring vs Winter - Combined Data") +
  theme(plot.title = element_text(hjust=0.5) )

ggsave("outputs/301/spring-winter-combined-data.jpeg")
```

Wowee! What's this mean?    
Running all chroms:
`bash $HOME/shernook/202-do-asso.sh $HOME/shernook/bamlists/bamlist156-winter-spring.bamlist $HOME/shernook/meta/GCF_002872995.1.chroms-less-test.chroms	 $HOME/shernook/bamlists/bamlist156-winter-spring.pheno`     
