---
title: "700-early-late-figures"
author: "Mac Campbell"
date: "January 12, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```


```{r}
library(tidyverse)
library(RColorBrewer)
```

Need to produce some figures for a paper.

__1__ Association testing
__2__ Haplo-raster of interesting bit?

### Association Testing
I do need to generate per chrom analyses of Early-Late, in outputs/300
```{sh, eval=FALSE}
bash $HOME/shernook/202-do-asso.sh $HOME/shernook/bamlists/bamlist156.bamlist $HOME/shernook/meta/GCF_002872995.1.chroms-less-test.chroms $HOME/shernook/bamlists/bamlist156-early-late.pheno
```
 
Spring/winter, in outputs/301
```{sh, eval=FALSE}
maccamp@farm:~/shernook/outputs/301$ for f in NC*lrt0.gz; do gunzip -c $f | grep -v "Chromosome" | awk '($6 > 2)' >> spring-winter.lrt0; done;
maccamp@farm:~/shernook/outputs/301$ gzip spring-winter.lrt0 
```
Filtering out lrt <=2 reduced file from 162 MB to 65 MB, but, let's focus on the main chromosomes
```{r}
#d1 <-  read_tsv("outputs/300/NC_037097.1-asso.lrt0.gz")
d16 <- read_tsv("outputs/300/NC_037112.1-asso.lrt0.gz")
d28 <- read_tsv("outputs/300/NC_037124.1-asso.lrt0.gz")

data<-bind_rows(d16,d28) 

data <- data %>% mutate(log10p = -log10(dchisq(LRT, df = 1))) %>%
  filter(log10p > 1) %>% filter(log10p!="Inf")

data %>% arrange(-log10p) %>% head()

```

```{r}
dd <- data %>% mutate(Index=1:n())

chroms<-dd %>% group_by(Chromosome) %>% mutate(Start=min(Index), Stop=max(Index)) %>% 
  select(Chromosome,Start,Stop) %>% unique() %>% 
  mutate(Position=round((Start+Stop)/2))

outliers <- dd %>% filter(log10p >=10)
#Make enough colors
nb.cols <- nrow(chroms)
mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(nb.cols)


plot1<-ggplot(dd) +
  geom_point(data=dd, aes(x=Index, y=log10p, color=Chromosome), alpha=0.75, cex=0.5) +
  geom_point(data=outliers, aes(x=Index, y=log10p), color="black", cex=0.5, alpha=0.9) +
  geom_hline(yintercept = 6, lty=2, alpha=0.6) +
  geom_hline(yintercept = 10, lty=2, alpha=0.6) +
  theme_bw() +
  scale_x_continuous(breaks=chroms$Position, labels=chroms$Chromosome) +
  theme(axis.text.x = element_text(angle=45, hjust = 1, face="bold", size=6)) +
  theme(legend.title = element_blank()) +
  theme(panel.grid = element_blank()) +
  theme(legend.position = "none") +
  scale_color_manual(values=mycolors) +
  ylab("-log10(p)") +
  xlab("Chromosome") +
  ggtitle("Early - Late Comparison") +
  theme(plot.title = element_text(hjust=0.5, face="bold", size=18) ) +
  theme(axis.text = element_text(size=12)) +
  theme(axis.title = element_text(size=14, face ="bold"))
```


```{r, eval=FALSE}
plot1
ggsave("outputs/700/association-early-late.jpeg",
       width=8, height = 5)
```
#### Set up for facet grid
```{r}
earlylate<-dd %>% mutate(Comparison="Early - Late")
```

## Spring-Winter
```{r}
#d1 <-  read_tsv("outputs/301/NC_037097.1-asso.lrt0.gz")
d16 <- read_tsv("outputs/301/NC_037112.1-asso.lrt0.gz")
d28 <- read_tsv("outputs/301/NC_037124.1-asso.lrt0.gz")

data<-bind_rows(d16,d28) 

data <- data %>% mutate(log10p = -log10(dchisq(LRT, df = 1))) %>%
  filter(log10p > 1) %>% filter(log10p!="Inf")

data %>% arrange(-log10p) %>% head()

```

```{r}
dd <- data %>% mutate(Index=1:n())

chroms<-dd %>% group_by(Chromosome) %>% mutate(Start=min(Index), Stop=max(Index)) %>% 
  select(Chromosome,Start,Stop) %>% unique() %>% 
  mutate(Position=round((Start+Stop)/2))

outliers <- dd %>% filter(log10p >=10)
#Make enough colors
nb.cols <- nrow(chroms)
mycolors <- colorRampPalette(brewer.pal(8, "Set1"))(nb.cols)


plot1<-ggplot(dd) +
  geom_point(data=dd, aes(x=Index, y=log10p, color=Chromosome), alpha=0.75, cex=0.5) +
  geom_point(data=outliers, aes(x=Index, y=log10p), color="black", cex=0.5, alpha=0.9) +
  geom_hline(yintercept = 6, lty=2, alpha=0.6) +
  geom_hline(yintercept = 10, lty=2, alpha=0.6) +
  theme_bw() +
  scale_x_continuous(breaks=chroms$Position, labels=chroms$Chromosome) +
  theme(axis.text.x = element_text(angle=45, hjust = 1, face="bold", size=6)) +
  theme(legend.title = element_blank()) +
  theme(panel.grid = element_blank()) +
  theme(legend.position = "none") +
  scale_color_manual(values=mycolors) +
  ylab("-log10(p)") +
  xlab("Chromosome") +
  ggtitle("Spring - Winter Comparison") +
  theme(plot.title = element_text(hjust=0.5, face="bold", size=18) ) +
  theme(axis.text = element_text(size=12)) +
  theme(axis.title = element_text(size=14, face ="bold"))


#### Set up for facet grid

springwinter<-dd %>% mutate(Comparison="Spring - Winter")

```


```{r, eval=FALSE}
plot1
ggsave("outputs/700/association-spring-winter.jpeg",
       width=8, height = 5)
```



#Combine
```{r}
df<-bind_rows(earlylate, springwinter)
```


```{r}
df$Comparison<-factor(df$Comparison, levels =c("Early - Late", "Spring - Winter"))
outliers <- df %>% filter(log10p >=10)

plot<-ggplot(df) +
  geom_point(data=df, aes(x=Position, y=log10p, color=Chromosome), alpha=0.75, cex=0.5) +
  geom_point(data=outliers, aes(x=Position, y=log10p), color="black", cex=0.5, alpha=0.9) +
  geom_hline(yintercept = 6, lty=2, alpha=0.6) +
  geom_hline(yintercept = 10, lty=2, alpha=0.6) +
  theme_bw() +
  theme(axis.text.x = element_text(angle=45, hjust = 1, face="bold", size=6)) +
  theme(legend.title = element_blank()) +
  theme(panel.grid = element_blank()) +
  theme(legend.position = "none") +
  scale_color_manual(values=mycolors) +
  ylab("-log10(p)") +
  xlab("Chromosome") +
  ggtitle("Association Testing") +
  theme(plot.title = element_text(hjust=0.5, face="bold", size=18) ) +
  theme(axis.text = element_text(size=12)) +
  theme(axis.title = element_text(size=14, face ="bold")) +
  facet_grid(Comparison~Chromosome, scales="free")
```


```{r}
plot
ggsave("outputs/700/association.jpeg",
       width=8, height = 5)
```
